\ttitle

### 2015-08-17 Chapters/CPET_Jaundice.tex
CPET Jaundice - Major update
- Next few commits will be following Donny chat this morning

\section{Results}
One-hundred and thirty eight patients had undergone pancreaticoduodenectomy (n=138), with preoperative cardiopulmonary exercise testing during the study period. Over half the patients were male (n=93, 67\%). Approximately half the number of patients were over the age of 65 (n=68, 49\%) and overweight or obese (n=69, 50\%). Cardiovascular comorbidity was present in 58 patients (42\%) and respiratory comorbidity was present in 12 patients (9\%). Fifty patients (36\%) had a history of cigarette smoking. The POSSUM Physiology Score was greater than 14 in 61 patients (44\%). Obstructive jaundice (serum bilirubin 35 – 250) was present in 32 (23\%) patients while severe obstructive jaundice (serum bilirubin > 250) was present in 19 (14\%) patients. 
The baseline demographic and clinical characteristics of non-jaundiced and jaundiced patients are shown in Table \ref{table:cpet_oj_patient}. A larger proportion of jaundiced patients were females compared to the non-jaundiced cohort (p$<$0.05) and smokers (p$<$0.05). Patients with jaundice were more likely to have an elevated POSSUM Physiology Score (p$<$0.005). Patients with cancer were more likely to be jaundiced (p$<$0.001). However, there was no statistically significant difference in age, BMI, cardiovascular comorbidity, or respiratory comorbidity between the non-jaundiced and jaundiced patients.

@@

The relationship between obstructive jaundice and preoperative blood tests is shown in Table \ref{table:cpet_oj_bloods}. While obstructive jaundice was statistically associated with multiple haematological and biochemical abnormalities, most of these did not appear to be of clinically significance. As expected, obstructive jaundice was associated with markedly elevated liver enzymes with severity of derangement associated with severity of jaundice. Obstructive jaundice and sOJ were associated with increasing CRP levels (p$<$0.001) and decreasing serum albumin levels (p$<$0.001). Obstructive jaundice was not associated with deranged renal function with both urea and creatinine remaining similar across all cohorts (p=0.09 and p=0.22 respectively).

@@

There was no association between obstructive jaundice and preoperative pulmonary function tests (Table \ref{table:cpet_oj_pft}).

@@

\subsubsection{Univariate analysis of obstructive jaundice versus CPET}
The relationship between obstructive jaundice and multiple physiological parameters measured at cardiopulmonary exercise testing is shown in Table \ref{table:cpet_oj_anaerobic}. There was an inverse relationship between oxygen consumption at the anaerobic threshold ($\dot{V}_{O_2}$AT) and increasing severity of jaundice (p$<$0.05). However, no such linear relationship was noted between any of the other parameters measured both at anaerobic threshold and at peak exercise in spite of apparent statistically significant associations.

@@

\subsubsection{Association between preoperative clinico-pathological factors and $\dot{V}_{O_2}$AT}
On multivariate analysis female sex (HR 3.75 CI 1.57-8.95 p$<$0.005), high BMI (HR 3.65 CI 1.61-8.26 p$<$0.005), presence of cancer (HR 4.02 CI 1.33-12.16 p$<$0.05) and raised CRP (HR 2.98 CI 1.29-6.86 p$<$0.05) were independently associated with low $\dot{V}_{O_2}$AT ($<$10mls/kg/min). However, jaundice was not associated with low $\dot{V}_{O_2}$AT. These results are shown in Table \ref{table:cpet_oj_at_regression}

@@

\subsubsection{Scatter-plot analysis}
Scatter-plot analysis comparing serum bilirubin and $\dot{V}_{O_2}$AT as continuous variables is depicted in Figure 1. This shows that the relationship between serum bilirubin and AT is weak with an r2 value of only 0.04 (I will have to confirm this but it is not more than 0.1).

>>>>

\section{Results}
One-hundred and thirty eight patients underwent pancreaticoduodenectomy with preoperative cardiopulmonary exercise testing during the study period. Over half of the patients were male (n=93, 67\%). Approximately half the number of patients were over the age of 65 (n=68, 49\%) and overweight or obese (n=69, 50\%). Cardiovascular comorbidity was present in 58 patients (42\%) and respiratory comorbidity was present in 12 patients (9\%). Fifty patients (36\%) had a history of cigarette smoking. The POSSUM Physiology Score was greater than 14 in 61 patients (44\%). Obstructive jaundice (serum bilirubin 35 – 250) was present in 32 (23\%) patients while severe obstructive jaundice (serum bilirubin $>$ 250) was present in 19 (14\%) patients.

The baseline demographic and clinical characteristics of non-jaundiced and jaundiced patients are shown in Table \ref{table:cpet_oj_patient}. A greater proportion of jaundiced patients were females compared to the non-jaundiced cohort (p=0.028) and smokers (p=0.038). Elevated POSSUM Physiology Score (p=0.004) and malignancy (p$<$0.001) were significantly associated with the presence of jaundice. However, there was no statistically significant difference in age, body mass index, cardiovascular comorbidity, respiratory comorbidity or preoperative biliary drainage between the non-jaundiced and jaundiced patients.

@@

The relationship between obstructive jaundice and preoperative blood tests is shown in Table \ref{table:cpet_oj_bloods}. Obstructive jaundice was associated with the presence of a systemic inflammatory response as characterised by raised serum C-reactive protein levels and decrease in serum albumin levels (p$<$0.001). The degree of systemic inflammation was proportionate to the severity of jaundice. Non-jaundiced patients had a median CRP of 3.6 mg/dl (inter-quartile range 0.3 - 89) and serum albumin of 37 g/dl (IQR 18 - 46) while patients with severe obstructive jaundice had a median CRP of 13 (IQR 1.7 - 51) and a median serum albumin of 25 g/dl (IQR 18 - 33). Obstructive jaundice was also associated with electrolyte abnormalities including serum sodium, serum potassium and serum chloride levels. Jaundiced patients were more likely to be anaemic (p$<$0.001) with lower haematocrit (p$<$0.001) and lower mean corpuscular volume (p=0.001). However, there was no significant difference in the preoperative renal function, prothrombin time and white cell count between jaundiced and non-jaundiced patients.

@@

Pulmonary function tests including forced vital capacity (FVC), forced expiratory volume in one second (FEV1) and their derived parameters (predicted FEV1, FEV1/FVC, predicted FEV1/FVC) were compared between jaundiced and non-jaundiced patients. There was a trend towards lower forced vital capacity with increasing severity of jaundice but this did not reach statistical significance (p=0.092). There was no association between the other pulmonary function tests and jaundice. (Table \ref{table:cpet_oj_pft}).

@@

Cardiopulmonary exercise test parameters measured at the anaerobic threshold and at peak exercise were compared between jaundiced and non-jaundiced patients using the Kruskal-Wallis test. Obstructive jaundice was associated with lower tidal volume (p=0.017), lower absolute $\dot{V}_{O_2}$ (p=0.003), lower corrected $\dot{V}_{O_2}$ (p=0.029), lower oxygen pulse (p=0.037) and lower respiratory rate (p=0.022). At peak exercise, jaundice was associated with lower tidal volume (p=0.028), lower absolute $\dot{V}_{O_2}$Peak (p=0.007), lower absolute $\dot{V}_{CO_2}$ (p=0.016) and lower end-tidal $O_2$ or $PET_{O_2}$ (p=0.026). However, as shown in Tables \ref{table:cpet_oj_anaerobic} and \ref{table:cpet_oj_peak}, several of these associations were not linear with a trend towards poorer results between non-jaundiced and mildly jaundiced patients and better values in the severely jaundiced cohort.

@@

Binary logistic regression analysis was undertaken to assess preoperative clinico-pathological patient factors associated with a low $\dot{V}_{O_2}$AT and low $\dot{V}_{O_2}$Peak. On univariate analysis, female sex (HR 2.74, 95\% CI 1.30-5.74, p=0.008), body mass index $>$ 25 (HR 3.09, 95\% CI 1.51-6.32, p=0.002), presence of malignancy (HR 3.59, 95\% CI 1.36-9.43, p=0.010), POSSUM Physiology Score $>$ 14 (HR 2.06, 95\% CI 1.02-4.17, p= 0.044), serum bilirubin $>$ 250 $\mu$mol/l (HR 5.66, 95\% CI 1.87-17.16, p=0.002), haemoglobin $<$ 12 g/dl (HR 2.74, 95\% CI 1.30-5.74, p=0.008) and serum C-reactive protein $>$ 10 mg/dl (HR 2.18, 95\% CI 1.06-4.51, p=0.035) were associated with $\dot{V}_{O_2}$AT $<$ 10 ml/kg/min. Age, cardiovascular and respiratory comorbidity, preoperative biliary drainage, mild jaundice and serum albumin were not significantly related to a low $\dot{V}_{O_2}$AT.

The statistically significant independent variables with p$<$0.05 were entered into a backward step-wise regression model with $\dot{V}_{O_2}$AT $<$ 10 ml/kg/min as a categorical, binomial, dependent variable. Female sex (HR 3.75, 95\% CI 1.57-8.95 p$<$0.005), body mass index $>$ 25 (HR 3.65, 95\% CI 1.61-8.26, p$<$0.005), malignancy (HR 4.02, 95\% CI 1.33-12.16, p$<$0.05) and C-reactive protein $>$ 10 mg/dl (HR 2.98, 95\% CI 1.29-6.86, p$<$0.05) were independently related to low $\dot{V}_{O_2}$AT. Obstructive jaundice was not related to low $\dot{V}_{O_2}$AT on multivariate analysis. (Table \ref{table:cpet_oj_at_regression})

Binary logistic regression analysis of the relationship between patient factors and low $\dot{V}_{O_2}$Peak ($<$ 16 ml/kg/min) is shown in Table \ref{table:cpet_oj_peak_regression}. On both univariate and multivariate analysis, female sex, body mass index $>$ 25 and haemoglobin $<$ 12 g/dl were significantly associated with a low $\dot{V}_{O_2}$Peak.

@@

Scatter-plot analysis comparing serum bilirubin versus $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak as continuous variables is depicted in Figure \ref{fig:cpet_oj_scatter}. This shows that the relationship between serum bilirubin and $\dot{V}_{O_2}$AT is weak with a $\rho^2$ value of 0.035 Pearson's $\rho$ = - 0.187, p = 0.028. There was no correlation between serum bilirubin and $\dot{V}_{O_2}$Peak (Pearson's $\rho$ = - 0.132, $\rho^2$ = 0.017, p = 0.123)

\begin{figure}[p]
	\centering
	\begin{subfigure}{0.6\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/cpet_oj_scatter_at_bil}
		\caption{Pearson's $\rho$ = - 0.187, $\rho^2$ = 0.035, p = 0.028}
		\label{fig:cpet_oj_scatter_at_bil}
	\end{subfigure}
	\begin{subfigure}{0.6\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/cpet_oj_scatter_peak_bil}
		\caption{$\rho$ = - 0.132, $\rho^2$ = 0.017, p = 0.123}
		\label{fig:cpet_oj_scatter_peak_bil}
	\end{subfigure}

	\caption{Pearson's correlation and scatter-plot analysis comparing serum bilirubin versus $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak in patients undergoing pancreaticoduodenectomy.}
	\label{fig:cpet_oj_scatter}

\end{figure}
